---
title: "Getting started with enact"
vignette: >
  %\VignetteIndexEntry{Getting started with enact}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(enact)
library(enfold)
library(rticles)
set.seed(443)
```

## Overview

`enact` is a package for estimating the effects of interventions via TMLE.
What differentiates it from other packages is its composability: it follows
your study from start to finish, handling the full implementation of a complex study
protocol and bundling all results in one neat object. 

It is also designed to be granular. Rather than providing monolithic functions that require you
to specify your entire study from start to finish at call time, you can
solve small tasks and review before moving on. This mirrors how carrying out studies
with a study protocol is approached. For example, if you wanted to investigate
estimated clever covariates, you can do so before estimating target parameters 
and peeking at results. This design has a few advantages:

- **Common metadata.** For example, a label specified for a variable is stored on the task, and
creating table 1's or plots will always know what you want your variables to be named.
Rather than interfacing packages for common reporting tasks and outcomes repeatedly,
`enact` interfaces them for you.
- **Flexible intervention specification.** `enact` requires you to specify interventions
explicitly, and has support for complex interventions. Modified treatment policies can
be specified and are estimated via a density classification trick.
- **Flexible learning.** `enact` uses `enfold` for learning tasks, which is a modern
model ensembling package boasting many features. You can send learners to different
treatment or outcome columns, giving you full control over how nuisance models are
estimated.

This vignette will introduce a basic workflow via `enact`.

## Running example

Suppose we followed a cohort of rheumatoid arthritis patients that were treated with
a Janus Kinase inhibitor (JAKi), a Tumor Necrosis Factor inhibitor (TNFi), or some
control treatment. We want to estimate risk differences and relative risks for
mortality and herpes zoster incidence, adjusting for age, sex, herpes zoster history,
and a disease score. Let us take a look at the (simulated) data:

```{r data}
data("enact_demo")
head(enact_demo)

nrow(enact_demo)
```

Looks straightforward, with one complication: the herpes zoster outcome has some missings.
The variable \code{censoring_hz} indicates missing outcomes. We will handle this downstream.

## Setting up a study

To set up a study, we use `initiate_study`. We set `verbose = TRUE`,
which prints useful information when extending the task.

```{r init}
my_enact_study <- initiate_study(
  data = enact_demo,
  confounders = c("age", "sex", "hz_history", "disease_score"),
  confounder_labels = c("Age", "Sex", "HZ history", "Disease score"),
  verbose = TRUE,
  cluster = NULL
)
my_enact_study
```

We can now tell the task what are treatment and outcome variables are.
We have two of each, so we add two of each and add labels for reporting.
Since the herpes zoster outcome has censoring, we tell the task which column
has information on censoring. The convention is `0` for censored observations,
and `1` for uncensored observations. `contrasts` tells the task which
target parameters we are looking to estimate.

```{r tx and outcome}
my_enact_study <- my_enact_study |>
  add_treatment(which = "tx_jaki", label = "JAKi") |>
  add_treatment(which = "tx_tnfi", label = "TNF inhibitor") |>
  add_outcome(
    which = "outcome_mortality",
    label = "Mortality",
    contrasts = list(ate(), log_relative_ate())
  ) |>
  add_outcome(
    which = "outcome_hz",
    label = "Herpes zoster",
    censoring = "censoring_hz",
    contrasts = list(ate(), log_relative_ate())
  )
```

In `add_outcome()`, we can also specify adjustment sets on a
per-outcome basis, if we wanted.

We can now also create a Table 1. We tell `enact` to stratify 
by treatment groups. Under the hood, `create_table_one` uses
the `gtsummary` package.

```{r table one}
my_enact_study <- my_enact_study |>
  create_table_one(stratify = c("tx_jaki", "tx_tnfi"))

my_enact_study$table_one
```

## Defining interventions

We will add four interventions. Three of them are just:
set `tx_jaki` and `tx_tnfi` to zero (everyone receives control),
or set either of them to one (everyone receives a JAKi, everyone receives a TNFi).
The last is a stochastic treatment that treats with a JAKi half the time,
and otherwise assigns control.

The stochastic intervention requires us to specify a custom intervention.
We can do so via `intervention()`. This is how it works:

```{r custom intervention}
my_intervention <- intervention(
  intervene = function(a, l) {
    a[] <- 0
    a$tx_tnfi <- rbinom(nrow(a), 1, 0.5)
    a
  },
  mtp = TRUE,
  label = "Stochastic"
)
```

`intervene` is a function that gives us our treatment variables and our global confounder
sets and lets us meddle with them however. We just need to remember to return the modified
treatment data in `a` at the end. The `mtp` flag specifies whether the density ratio
classification trick should be used. This is especially useful for modified treatment
policies (hence the argument name), but applicable whenever the natural and intervened-on
treatment distributions share the same support (?).

We now add this intervention to our task, as well as a few intervention templates.
`static_intervention` sets all treatments to the same value, and `intervention_arm`
sets the specified column to `1` and others to `0`. The `reference` argument
tells the task which intervention to treat as the reference when formatting results.

```{r interventions}
my_enact_study <- my_enact_study |>
  define_interventions(
    my_intervention,
    static_intervention(0, label = "Control"),
    intervention_arm("tx_jaki", label = "JAKi"),
    intervention_arm("tx_tnfi", "TNFi"),
    reference = "Control"
  )
```

## Defining learners via enfold

We will now add specify some learners. This is functionality of the
`enfold` package, and you are encouraged to read its vignette to
understand how it works. We will provide a basic ensemble here:
a mean learner, a generalized linear model, a generalized additive model,
Multivariate Adaptive Regression Splines, and gradient boosting.
For gradient boosting, we use the list learner concept in `enfold`
to fit the model once and get results after 10, 20, ..., 100 iterations.
We also specify a superlearner as our metalearner.

```{r learners}
my_mean <- lrn_mean("Mean")
my_glm <- lrn_glm("GLM", family = binomial())
my_gam <- lrn_gam("GAM", binomial())
my_xgb <- lrn_xgboost(
  "XGB",
  objective = "reg:logistic",
  eta = 0.1,
  nrounds = 100L,
  n_snapshots = 10L
)

my_learners <- list(my_mean, my_glm, my_gam, my_xgb)
my_superlearner <- mtl_superlearner("SL", loss_fun = loss_logistic())
```

We can now add them to the task. Since we have an intervention with `mtp = TRUE`
we need to provide learners for those interventions that can output classification probabilities
for a binary outcome. The same applies to censoring. The selector functions
`treatments()` and `outcomes()` have an optional argument named `which` that
we can use to assign particular learner-metalearner combinations to individual treatments
or outcomes via their labels. We do not use this functionality here.

Before adding learners, we also need to specify cross-validation folds. We use
`inner_cv = 2L` and `outer_cv = 2L` to decrease runtime. You probably want to increase
these for a real study. `inner_cv` controls how many folds are used to build ensembles.
`outer_cv` controls how many ensembles are built, and if it is not `NULL`
cross-validated predictions are estimated. **This is necessary for many
learners when using TMLE**. If our task has a cluster variable, cross-validation
is clustered automatically.

```{r add learners}
my_enact_study <- my_enact_study |>
  add_cv_folds(inner_cv = 2L, outer_cv = 2L) |>
  add_models(
    treatments(learners = my_learners, metalearner = my_superlearner),
    outcomes(learners = my_learners, metalearner = my_superlearner),
    mtp(learners = my_learners, metalearner = my_superlearner),
    censoring(learners = my_learners, metalearner = my_superlearner)
  )
```

## Fitting interventions and checking balance

Everything is specified now and we can fit! `enact` explicitly separates
fitting interventions and outcomes. We make use of this to perform some
balance checks on our intervention (treatment and censoring) models.

For this purpose, we use the `add_balance_checks()` function. Using
`check_balance()` then prints a balance check summary, but the task also
stores love plots and balance tasks. These are included in the final report
(see below), so you don't have to extract them just yet, if you don't want to.

```{r fit}
my_enact_study <- suppressWarnings(
  my_enact_study |>
    fit_interventions() |>
    add_balance_checks() |>
    check_balance()
)
```

There are many violations here, which might be a good sign to investigate
our learners in more detail or change estimation strategies. We'll let this
pass for the vignette. However, notice how we were able to detect the
balance issue here *before calculating any results*. In a real study,
we could fix this problem now without peeking.

(It seems the balance checks didn't print for the reference intervention.
That's... a bug, right?)

## Estimate results via TMLE

Almost there! We now also fit outcomes and then target all nuisance estimates
via TMLE. For this, we use `do_tmle()`. Since all our outcomes are binary,
we can just tell the task to target all outcomes via a binomial fluctuation.

Note that `enact` estimates standard errors for TMLE via a fractional weighted
(Bayesian) bootstrap of the targeting step. In some simulation studies, this
has markedly better (frequentist) properties than using the variance of the
efficient influence function. If we have clusters, the bootstrap is clustered
at the cluster level.

After targeting, our task object provides TMLE results via `summary()`.

```{r tmle}
my_enact_study <- suppressWarnings(
  my_enact_study |>
    fit_outcomes() |>
    do_tmle(
      which = all_outcomes(),
      fluctuation_family = binomial()
    )
)

summary(my_enact_study)
```

## Strip and report

We have successfully estimated everything! At this point, there
are two more functions to know of. 

The first one is `strip()`. Calling `strip()` on a task object
removes everything from it that is not a result. This is helpful in cases
where you want to send results to colleagues for reporting,
but can't send the full task for data privacy reasons. By default,
`strip()` keeps table 1's, balance checks, and TMLE results.

After calling `strip()`, we can call `report()` on the stripped object.
Calling `report()` creates a .docx file for Word or a .tex/.zip file for
LaTeX. For the latter, one can use templates from the `rticles` package.
This report can be used as a basis for a manuscript, or to copy paste results
cleanly. The code below puts this report in a temporary directory.
It can't be shown in the vignette, but if you re-run it you can investigate
the results in that temporary folder!

```{r report}
report_dir <- tempfile(fileext = ".zip")

my_enact_study <- my_enact_study |>
  strip()

make_report <- report(
  task = my_enact_study,
  output_file = report_dir,
  format = "tex",
  template = rticles::sage_article(),
  include_love_plots = TRUE,
  include_density_plots = TRUE
)
```